---
title: "Flattering Advice: Avoiding Disappointment as a Driver of Gender Discrimination"
shorttitle        : "Flattering Advice"
author: 
  - name          : "Amanda Chen"
    affiliation   : "1"
    corresponding : yes    
    email         : "zchengj@connect.ust.hk"
    address       : "Department of Management, The Hong Kong University of Science and Technology, Hong Kong"
  - name          : "David Hagmann"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "The Hong Kong University of Science and Technology"
note : |
  `r format(Sys.time(), "%e %B, %Y")`
authornote : |
  The authors are grateful to a lot of things.
abstract: |
  The purpose of advice is to help recipients make better decisions and solve their problems. However, in this paper, we propose that advisors may also take recipients' responses into account and attempt to avoid disappointing them, leading to what we call "flattering advice." In two pre-registered experiments involving real interactions between advisors and advisees (n = 2,700), we show that individuals consider others' expectations when giving advice, even when such expectations do not provide informative insights for the underlying decision-making problem. As a result, men receive more aspirational advice than women, given their higher confidence, when expectations (but not gender) are shown to advisors. However, incorporating expectations leads to worse outcomes (Study 1). Individuals inflate their advice due to interpersonal considerations: people who are concerned about advisees' likability inflate their advice to a great extent. Advisees reward such flattering advice, finding the advisers more likable and no less trustworthy than those who provide more accurate advice (Study 2). We discuss practical and theoretical contributions and future research directions.
 
keywords          : "Gender difference, Advice, Interpresonal relationship"
wordcount         : "4,362"
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
header-includes:
   - \usepackage{setspace, appendix, placeins, booktabs, tikz, siunitx}
   - \usepackage{tabularx, epigraph, float, colortbl, tabu, pdflscape}
   - \usetikzlibrary{positioning}
   - \raggedbottom
   - \usepackage[maxfloats=256]{morefloats}
   - \maxdeadcycles=1000
bibliography      : bibliography.bib
output: 
  # papaja::apa6_docx
  papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
always_allow_html: true
---

```{r setup, include=FALSE}
library(magrittr)

knitr::opts_chunk$set(fig.width = 8, fig.height = 4, cache = F, echo = F,
                      fig.align = 'center',warning = FALSE, message = FALSE)

study1_math50 <- read.csv(here::here("Data/Cleaned Data/S1_Math50_cleaned.csv"))
study1_time1All <- readRDS(here::here("Data/Study1_time1All.Rds"))
study1_time1 <- readRDS(here::here("Data/Study1_time1.Rds"))
study1_time2_giver <- readRDS(here::here("Data/Study1_time2_giver.Rds"))
study1_time2_giver50 <- readRDS(here::here("Data/Study1_time2_giver50.Rds"))
study1_advice <- readRDS(here::here("Data/Study1_time2_advice.Rds"))
study1_time3 <- readRDS(here::here("Data/Study1_time3.Rds"))

```

# Introduction

Seeking and giving advice can present conversational challenges that
are difficult to navigate. Advice seekers may be embarrassed to ask for
help [@Benabou2022], and they may fail to do so if the information they
receive could lead to disappointment [@Gill2012]. Indeed, people prefer
advisers who withhold unpleasant information [@Shalvi2019] and punish
those who tell them things they do not want to hear [@John2019]. People
appear to engage in strategic selection of advice to avoid unpleasant
information and protect the utility they get from positive beliefs about
themselves and the future they anticipate [@Loewenstein2018;
@Loewenstein1987]. But do advice givers anticipate and consider the
hedonic cost of honest information to the recipient? While previous work
has emphasized the instrumental value of advice (e.g., satisfaction with
a decision or accuracy of a prediction), we focus on the motivations of
the person providing advice.

Research on advice has largely focused on settings that do not tackle
belief utility. For example, participants may have been asked to
estimate the number of balls in an urn or choose from pairs of lotteries
[@Benjamin2015]. Advisers have been shown to be motivated to give
accurate advice [@Jonas2003], and engage in perspective-taking to meet
the preferences of the advisee [@Kray2000; @Kray1999]. When they are
uncertain, they even prefer that advisees do not rely (extensively) on their
recommendations [@Ache2020]. Thus, it seems that people are concerned
about providing good advice even when they are not incentivized for the
recipient's decision quality.

In many real settings, however, advice incorporates beliefs about the
recipient. Someone recommending an expensive restaurant conveys that
they think the advisee is wealthy enough to afford an expensive meal,
and someone recommending a student pursue a PhD in economics assumes
that they are intellectually capable. Conversely, telling a poorly
performing student to switch to a less-demanding major conveys a belief
that the student is not capable of improving. Thus, while the painful truth
may prevent someone from making a bad decision, it may come at a hedonic cost to
the recipient and damage the relationship between the adviser and
advisee. We propose that advisers take these costs into account, even
when the interactions are entirely anonymous and they are incentivized
for the quality of the advisee's decision.

Specifically, we focus on the signaling value of advice: what does the
advice imply about the adviser's perception of the advisee? We propose
that advisees have expectations about the advice they receive and treat
this as a reference point [@Koszegi2006]. Advice that falls short of
these expectations creates disappointment, while advice that exceeds
them is flattering and creates positive interpersonal benefits because
it creates a perception that the adviser holds the advisee in high
esteem. Such interpersonal considerations may be as important, if not
more important, to the adviser than the instrumental value of the
advice: the adviser directly experiences a souring relationship, but may
suffer little and only much delayed consequences if an advisee makes a
poor life decision.

Our account therefore predicts that people give and receive advice that is more
flattering and suggestive of inflated ability and optimistic outcomes
than would advice in the absence of such interpersonal considerations.
However, we propose that this is not true for all advisees equally.
Presenting flattering advice requires the formation of beliefs about the
advisee's expectations (reference point). @Exley2022 show that men are
more overconfident than women, conditional on equal performance, and that
people anticipate this difference. An adviser who seeks to avoid disappointment and incorporates
expectations into their advice would then present more favorable advice
to male advisees than to female advisees (see Figure 1 for an
illustration). This provides a novel account for a well-known finding
that men receive more aspirational advice than do women [@Kanze2018].
Notably, when it pays to have accurate beliefs, such flattering advice
would come at a cost: while men would be more likely to aim higher, they
would also suffer greater costs from aiming too high and failing to meet
the objective.

In a large, preregistered experiment (n = 2,000), we show that advisers
consider the expectations of advisees, even when such expectations do
not provide informative insights for the underlying decision-making
problem. We find that as a result, showing expectations (but not gender)
to advisers leads men to receive more aspirational advice than women.
Incorporating expectations leads to worse advice, even when advisers are
incentivized based on the outcome of the advisee's decision. In a second
preregistered experiment (n = 700), we show that advisers respond to incentives for
interpersonal considerations. They tell people that they are more
attractive than they believe them to be, and inflate to a greater extent
when incentivized for likability rather than accuracy. Notably, they inflate
even when only incentivized for accuracy. Advisees indeed
reward such flattering advice, rating the advisers more likeable and no
less trustworthy than those who provide more accurate advice.

# Open Science Statement

We report all sample sizes, data exclusions, manipulations, and
measures in the studies. Screen captures of the experimental materials
are available in the Supplemental Information. The complete data, code
to reproduce all statistical analyses and figures in the manuscript, as
well as the preregistration reports are available via OSF. All our
studies were preregistered on AsPredicted.org.^[<https://osf.io/8r3d4/?view_only=5ad7bafcd16b4d4ba08bb28b0e2bd02d>]

# Study 1

We begin by examining whether advisers take the expectations of advisees
into account, even when those expectations are not relevant to the decision
the advisee has to make. In a three-stage experiment, we first recruit a
sample of advisees and ask them to complete a mathematics quiz. After
answering ten multiple-choice questions, participants are anchored
to low or high performance expectations and asked to guess how many
questions they answered correctly. Next, we recruit advisers and
show them either only the true performance of the advisee, or the true
performance and how many questions they guessed they answered correctly.
Advisers are tasked with recommending whether the advisee should compete
against a group of high-performers or a group of low-performers on the
mathematics quiz, where a potential bonus depends on outperforming the competitor
and which group was chosen. Finally, we invite the advisees back and show them the
advice they have received. Advisees then pick whether to compete against
high or low performers, with the outcome determined by their past performance on the quiz. Our
key hypotheses are that (1) advisees who were anchored to high expectations
would be more likely to receive advice to compete against the high
performer group, and (2) that men are more likely than women to be advised to 
compete against the high-performer group when expectations are shown.

## Methods

```{r}

subgroup_averages <- data.frame(subgroup = numeric(), avg_score = numeric())
set.seed(807)
# loop through 1000 iterations of randomly selecting 10 participants and computing their average score
for (i in 1:1000) {
  # randomly select 10 participants
  subgroup <- sample(study1_math50$ProlificID, size = 5, replace = FALSE)
  # compute the average score for the selected participants
  subgroup_avg <- mean(study1_math50$Score[study1_math50$ProlificID %in% subgroup])
  # add the subgroup and its average score to the dataframe
  subgroup_averages <- rbind(subgroup_averages, data.frame(subgroup = i, avg_score = subgroup_avg))
}

p5 <- quantile(subgroup_averages$avg_score, probs = 0.05)
p95 <- quantile(subgroup_averages$avg_score, probs = 0.95)

sortedScore <- sort(study1_math50$Score, decreasing = TRUE)
LP_lower <- sortedScore[50]
LP_upper <- sortedScore[31]
HP_lower <- sortedScore[20]
HP_upper <- sortedScore[1]

```

As part of our main experiment, participants choose to compete against a group
of previous participants and have their expectations anchored based on their performance.
Thus, we begin by first recruiting a sample of `r scales::comma(dim(study1_math50)[1])`
participants from Prolific to complete a 10-question multiple choice
mathematics quiz. The questions were taken from a paper-version of the
ASVAB standardized exam, such that answers were not available online.
Participants had five minutes to answer the quiz and were paid 10 cents for
each correctly answered question. We define the top 20 scorers as the
"High Performers" and the bottom 20 scorers as the "Low Performers." On
average, participants answered `r round(mean(study1_math50$Score),2)`
questions correctly, High Performers scored between `r HP_lower` and
`r HP_upper`, and Low Performers scored between `r LP_lower` and
`r LP_upper`. In order to anchor the expectations of participants in our main
experiment, we simulated 1,000 pairings of groups of five
participants, with the 5<sup>th</sup> percentile of groups scoring an average of
`r p5` and the 95 th percentile scoring an average of `r p95`. We report these
averages to participants in the Low Expectations and High Expectations treatment,
respectively.

We then recruited `r scales::comma(dim(study1_time1All)[1])`
participants for the role of advisees in Stage 1 of our main experiment.
To arrive at a gender-balanced sample, we dropped the last two
male participants to complete the survey, ending up with a sample of 500
men and 500 women ($M_{\text{Age}}$ = `r mean(study1_time1$Age)`). 
Participants completed the same 10-item mathematics quiz as the
earlier participants and were informed that their performance would
affect their bonus earnings in a follow-up stage to be conducted a few
days later. After completing the quiz, we informed them of the average
score of a group of five participants from the preliminary survey. We
randomly assigned them to learn about the 5<sup>th</sup> percentile of groups,
which scored `r p5`. ("Low Expectations" treatment) or the 95<sup>th</sup>
percentile of groups, which scored `r p95` ("High Expectations"
treatment). Participants then made a guess (unincentivized) about how
many questions they think they answered correctly. The survey concluded
with basic demographic questions.

We then recruited `r scales::comma(dim(study1_time2_giver)[1])`
participants for Stage 2, placing them in the role of advisers. We began
by informing them of the mathematics quiz that participants in the
preliminary study and Stage 1 had completed, and informed them of the
average score of all participants in the preliminary study. Advisers had to recommend
whether an advisee should compete against the Low Performers or the High
Performers (we used these terms in the survey). We anticipated that
being told to compete against High Performers is more flattering and hence being
told to compete against the Low Performers would be disappointing if one had
expected to do well.
Advisees would earn 30 cents if their score was equal to or higher than
a randomly selected member of the Low Performer group and they chose to
compete against this group, or 50 cents if they performed as well or
better than a randomly selected member of the High Performer group and they
chose to compete against that group. If
they scored lower, they would not receive any bonus.

Advisers were randomly assigned to one of two treatments. In the
"Baseline" treatment, they only observed the score of the advisee on the
mathematics quiz. In the "Expectation" treatment, they observed the
score as well as the advisee's guess for how many questions they
answered correctly. Note that since the outcome is determined only by
the past score on the quiz, the advisee's guess is immaterial to which
group they should compete against. Moreover, in neither treatment did they
receive any demographic information about their advisee. Participants
gave recommendations to 10 advisees, which unbeknownst to them were five
men and five women matched to have identical performance on the test.^[We made this decision to account for the possibility of gender differences in performance.]
They were informed that if their advice was shown to a participant who
returned for the follow-up survey, they would receive the identical
bonus as that participant. The survey then concluded with basic
demographic questions. 

Finally, we invited participants from Stage 1
back for the follow-up survey. Following our preregistration, we kept
the survey open for 7 days. In total,
`r scales::comma(dim(study1_time3)[1])` participants
(`r scales::comma(dim(study1_time3[study1_time3$Gender == 'Male',])[1])`
men,
`r scales::comma(dim(study1_time3[study1_time3$Gender == 'Female',])[1])`
women) returned. The brief survey reminded them of the task they
completed in Stage 1, informed them that other participants from
Prolific had observed their real score and given them advice against
which group to compete against, and finally were reminded them of how many
questions they guessed they had answered correctly. Importantly, they
were not informed of their true score or the score of the groups they
could compete against. Participants then observed the advice from a
randomly selected adviser and made their decision.

```{r}
library("papaja")
# Fit linear regression model
study1_advice <- study1_advice %>% dplyr::mutate(ExpShow = ifelse(Condition == 'E',1,0))

model1 <- lm(Guess ~ Gender + count, data = study1_time1)
model1_apa <- apa_print(model1)
```

## Results

We begin by examining the performance of the Stage 1 participants.
Because our treatment takes place after participants completed the
mathematics quiz, we would not expect a difference in performance across
the Low and High Expectations treatments. Indeed, the two groups scored
no different from one another
(`r round(mean(study1_time1[study1_time1$Condition == "Low",]$count),2)`
and
`r round(mean(study1_time1[study1_time1$Condition == "High",]$count),2)`
for the Low Expectations and High Expectations treatments, respectively,
`r papaja::apa_print(t.test(count ~ Condition, data = study1_time1, var.equal = T))$statistic`).
The expectations treatment did, however, affect how well they thought
they performed. Participants in the "Low Expectations" treatment guessed
a score of
`r round(mean(study1_time1[study1_time1$Condition == "Low",]$count),2)`
vs.
`r round(mean(study1_time1[study1_time1$Condition == "High",]$Guess),2)`
in the High Expectations treatment
(`r papaja::apa_print(t.test(Guess ~ Condition, data = study1_time1, var.equal = T))$statistic`),
showing that the manipulation was successful, albeit small. Contrary to our
expectations, we did find a gender difference in performance: men scored
`r round(mean(study1_time1[study1_time1$Gender == "Male",]$count),2)` on
average, while women scored
`r round(mean(study1_time1[study1_time1$Gender == "Female",]$count),2)`
(`r papaja::apa_print(t.test(count ~ Gender, data = study1_time1, var.equal = T))$statistic`).
Consistent with this difference, men thought they answered more questions 
correctly than did
women
(`r round(mean(study1_time1[study1_time1$Gender == "Male",]$Guess),2)`
vs
`r round(mean(study1_time1[study1_time1$Gender == "Female",]$Guess),2)`,
`r papaja::apa_print(t.test(Guess ~ Gender, data = study1_time1, var.equal = T))$statistic`).
This difference, however, does not affect the interpretation of our
findings, which will rely on an interaction of gender with an
experimental treatment for Stage 2 participants.

We define a measure of "overconfidence" as (Performance - Estimate). We observe
that women underestimate their performance by 0.76 points, while men do so by
only 0.27 points `r papaja::apa_print(t.test(overconfidence ~ Gender, data = study1_time1 %>% dplyr::mutate(overconfidence = Guess - count)))$full_result`.
Although neither gender is overconfident, men on average are more confident in
their performance than are women, as we had expected.

<!-- We conducted an OLS regression with the performance and gender of -->
<!-- participants and found that with the same performance, male participants -->
<!-- are more confident than their female counterparts -->
<!-- (`r model1_apa$estimate['GenderMale']`, -->
<!-- `r model1_apa$statistic['GenderMale']`). -->

<!-- David: I don't think we can do this regression, but defining overconfidence with a difference score and doing a t-test should be fine. -->

```{r study1regs}
study1_advice <- study1_advice %>% dplyr::mutate(Treatment = factor(Treatment, levels = c('Low', 'High')))

# David: No need to redefine a factor variable as a continuous number. Easier to just change the reference level.

modelsummary::modelsummary(list(`(1)` = lm(Advice ~ Treatment + Performance , data = study1_advice %>% dplyr::filter(ExpShow == 1)),
                                `(2)` = lm(Advice ~ ExpShow*Gender , data = study1_advice),
                                `(3)` = lm(ExpectedBonus ~ ExpShow*Gender, data = study1_advice)),
                           vcov = c(~GiverID, ~GiverID,~GiverID),
                            coef_map = c("TreatmentHigh" = "High Expectation",
                                         "Performance" = "Performance",
                                         "ExpShow" = "Expectation Shown",
                                        "GenderMale" = "Advisee Male",
                                        "ExpShow:GenderMale" = "Expectation x Male",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "Column 1 displays the advice given to compete against a High Performance group based on whether the advisees were primed with low or high expectations and the expectation level shown to the advisor. Column 2 displays the advice given to compete against a High Performance group based on the gender of the targets and the expectation level shown to the advisor. Column 3 displays the expected bonus of the advice received based on the gender of the targets and the expectation level shown to the advisor. All standard errors clustered at the level of the advisor.") %>%
  kableExtra::kable_styling(latex_option = "scale_down")

```

```{r study1genderdiff, fig.cap ="Advisers observed real performance and the participants' estimated performance, but not their gender. As a result of their higher expectations, men (blue) were advised to compete against the high performance group more often than women (red). The grey line shows advice absent expectations, which did not differ by gender."}
library(dplyr)
library(ggplot2)
library(magrittr)

ProbHG <- study1_advice %>%
  group_by(Condition, Performance, Gender) %>%
  dplyr::summarize(avg_hardGroup = mean(Advice))


df_exp <- subset(ProbHG, Condition == "E")
df_none <- subset(ProbHG, Condition == "None")

ggplot(df_exp, aes(x = Performance, y = avg_hardGroup, color = Gender)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"),se = FALSE) +
  geom_smooth(data = df_none, aes(x = Performance, y = avg_hardGroup),
              method = "glm", method.args = list(family = "binomial"), color = "grey", se = FALSE) +
  scale_color_manual(values = c("red", "#4169E1", "grey"),
                     labels = c("Female + Expectations", "Male + Expectations", "Expectation Hidden")) +
  scale_y_continuous(name = "P(Compete Against High Performance)",
                     labels = scales::percent,
                     expand = c(0, 0)) +
  scale_x_continuous(name = "Performance",
                     expand = c(0, 0),
                     breaks = seq(0, 10, 2)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(1.2, "lines"),
        legend.text = element_text(size = 12),
        plot.title = element_text(hjust = 0.5, size = 20))


```

<!-- David: Is there something different about the figure below or can we delete? -->
```{r, eval = F}
library(dplyr)
library(ggplot2)
library(magrittr)

ProbHG <- study1_advice %>%
  group_by(Condition, Performance, Gender) %>%
  dplyr::summarize(avg_hardGroup = mean(Advice))


df_exp <- subset(ProbHG, Condition == "E")
df_none <- subset(ProbHG, Condition == "None")

ggplot(df_exp, aes(x = Performance, y = avg_hardGroup, color = Gender)) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"),se = FALSE) +
  scale_color_manual(values = c("#FFA500", "#4169E1"),
                     labels = c("Female_withExpectation", "Male_withExpectation")) +
  scale_y_continuous(name = "P(Compete Against High Performance)") +
  scale_x_continuous(name = "Performance") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(1.2, "lines"),
        legend.text = element_text(size = 12),
        plot.title = element_text(hjust = 0.5, size = 20))


```

Next, we examine whether advisers took the advisee's expectations into
account. Column 1 of Table \@ref(tab:study1regs) reports a linear
probability model on advising to compete against the High Performer
group for advisers in the "Expectations" treatment, controlling for the
true performance of the advisee. Because each adviser
made ten recommendations, we cluster standard errors at the adviser
level. However, contrary to our expectations, we do not see a significant
effect of the expectations treatment. This may be because the induced difference
in expectations was too small.

However, recall that men were more confident in their performance than were women.
Our theory thus predicts that showing expectations should lead to more flattering
advice (i.e., a recommendation to compete against the High Performer group) for
men than for women. We report a linear probability model with advice to
compete against the high group as the outcome measure, and the advisee's
gender, whether expectations were shown to the adviser, and the
interaction of the two in Column 2 of Table \@ref(tab:study1regs). As
predicted, we find a significant interaction effect: men are more likely
to be advised to compete against the High Performers when expectations
are shown than when they are hidden ($p < 0.001$).
We show this result graphically in Figure \@ref(fig:study1genderdiff)).
<!-- David: Can you add a description of the figure here? -->

To examine the quality of advice, we computed the expected bonus
earnings for someone who followed the recommendations. For example, if an
adviser suggested competing against the High Performer group, we matched the
advisee against all 20 members of that group and determined how often their score
matched or exceeded that of the member. We then multiplied this number by the
respective bonus earnings (50 cents and 30 cents for the High and Low Performer 
group, respectively). To see if including expectations leads to worse advise
for men, we report a linear probability model with the experimental treatment
of the adviser, the gender of the advisee, and their interaction in Column 3
of Table \@ref(tab:study1regs). Displaying expectations led men to be advised
to compete against the High Performer group more often, and this advise turned
out to be bad: men receive worse advise than do women when expectations are 
displayed, but not in their absence.^[This analysis was not preregistered, and we note here that the reduction in expected earnings is small. However, it is interesting that expectations have a negative effect for men who underestimate their performance on average. One possibility is that advisers suggest the High Performer group more often than is optimal. We return to the possibility that advice is overall biased to be flattering in Study 2.]

```{r}
study1_time3 <- study1_time3 %>% dplyr::mutate(AdviseeConditionHigh = ifelse(AdviseeCondition =='High',1,0))

model2 <- lm(bonus ~ AdviseeConditionHigh + AdviseeScore, data = study1_time3)
model2_apa <- apa_print(model2)

model3 <- lm(Bet ~ Gender + AdviseeScore, data = study1_time3)
model3_apa <- apa_print(model3)

```

To determine whether flattering advice is truly costly, however, we need
to examine the outcome of the advisees. In particular, they could ignore flattering
advice, recognizing it as such and thus failing to adhere to it.

<!-- David: Let's turn the regression you discussed below into a column of the Table above -->
<!-- David: Also, one more suggestion for an analysis: let's look at whether Advisees followed the recommendation they received as an other linear probability model (1/0). I would suggest predictors of whether it's the High Performer group, whether the adviser saw expectations, the gender of the advisee. Also, another model that has the two-way interaction for all of those. We may not end up reporting this, but I would be curious to see. -->

<!-- Finally, we look at the actual earnings of advisees. Although -->
<!-- participants in the High Expectation condition earned less, this -->
<!-- difference is not significant -->
<!-- (`r model2_apa$estimate['AdviseeConditionHigh']`, -->
<!-- `r model2_apa$statistic['AdviseeConditionHigh']`). -->

<!-- We found that males are more competitive to the extent that they are -->
<!-- more likely to choose to compete against High Performer group -->
<!-- (`r model3_apa$estimate['GenderMale']`, -->
<!-- `r model3_apa$statistic['GenderMale']`), holding performance the same. -->

## Discussion


In Study 1, we discovered that although advisers are encouraged to
provide accurate advice, they still take into account the expectations
of advisees, which are irrelevant to the utility of the advice. Due to
the fact that males tend to be more confident than their equally
competent female counterparts, they are more likely to be recommended to
compete against high-performance competitors. However, when advice is
given to match the advisees' expectations, these flattering
recommendations result in lower expected bonus earnings for male
participants.

# Study 2

We then conduct a three-stage experiment to determine why advisers give
flattering advice. First, we ask a group of advisees to upload selfies
and predict their rank in a group of ten people of the same gender based
on physical attractiveness. Next, we recruit and ask advisers to rank
selfies of the opposite gender and recommend which rank the participant
ranked 7th should bet on, with the bonus being contingent on whether the
participant guessed accurately or found the adviser likable. Finally, we
show the advisees the advice they have received and ask them to make
their final prediction about their rank and evaluate the advice giver in
terms of likability, warmth, and trustworthiness. Our main hypothesis is
that advisers who want to be liked by advisees would recommend a higher
rank.

```{r}

study2_time1 <- readRDS(here::here("Data/Study2_time1.Rds"))
study2_time1_200 <- readRDS(here::here("Data/Study2_time1_200.Rds"))
study2_time2 <- readRDS(here::here("Data/Study2_time2.Rds"))
study2_time3 <- readRDS(here::here("Data/Study2_time3.Rds"))

```

## Methods

We recruited 300 participants and invited them to upload photos of
themselves (selfies) to be rated by other participants on
attractiveness. We ended up with a sample of
`r scales::comma(dim(study2_time1[study2_time1$gender == 'Male',])[1])`
men and
`r scales::comma(dim(study2_time1[study2_time1$gender == 'Female',])[1])`
women who agreed to do so and uploaded usable pictures. We selected the
first 100 photos from women to arrive at a gender-balanced sample
($M_{\text{Age}}$ = `r mean(study2_time1_200$age)` years; 50.0% Female).
We informed them that their selfies will be randomly grouped into nine
others from the same gender and ranked by people of the opposite gender
and asked them to predict their rank, with 1st rank represents the most
attractive and 10th rank represents the least attractive. We assigned
each participant to a group of 10 participants of their own gender.

Next, we recruited `r scales::comma(dim(study2_time2)[1])`new
participants and asked them to rank the attractiveness of one such group
of 10 participants of the opposite gender ($M_{\text{Age}}$ =
`r mean(study2_time2$age)` years;
`r round(mean(study2_time2$gender == "Female")*100, 2)`% Female). After
they did so, we showed them the photo of the participant they ranked as
the 7th most attractive (that is, the 4th least attractive) and reminded
them that they had ranked this person at 7th attractive. We (truthfully)
informed them that this participant would have a chance to earn a bonus
if they accurately guessed their rank based on the aggregated ratings of
all participants who had ranked this group. Participants were then
randomly assigned to an "accuracy" or a "likeability" treatment. In the
accuracy treatment, they were told that they would receive a bonus (\$1)
identical to that of the advisee. While in the "likeability" treatment,
the amount their bonus depended on how the advisee rated them in terms
of likeability, after observing only the rank that they had advised the
advisee to guess. Advisees would rate to what extent they think the
adviser is likable on a 5-point scale, and each point they earn
increases their bonus by 20 cents.

Finally, we invited participants from Stage 1 and were ranked as the 7th
attractive by at least one adviser (so that they received advice) back
for the follow-up survey. Following our preregistration, we kept the
survey open for 7 days. In total,
`r scales::comma(dim(study2_time3)[1])` participants
(`r scales::comma(dim(study2_time3[study2_time3$gender == 'Male',])[1])`
men,
`r scales::comma(dim(study2_time3[study2_time3$gender == 'Female',])[1])`
women) returned. They were reminded of the selfie they uploaded in Stage
1 and informed that a group of 10 selfies, including theirs, had been
rated by other participants from Prolific. They were also introduced to
the opportunity to earn a \$1 bonus if they guessed their rank
accurately. Participants were then shown the advice from a randomly
selected adviser and made their predictions. After deciding which rank
to bet on, advisees were reminded of the advice they received and asked
to rate the advice giver's likability, warmth, friendliness,
good-naturedness, trustworthiness, and sincerity on a 5-point scale. The
last 6 items were adopted from @Fiske2007.

## Results

```{r study2advice, fig.cap="Advice for 7th most attractive participants"}
library(dplyr)
library(ggplot2)

count <- study2_time2 %>%
  group_by(Condition) %>%
  summarize(number = dplyr::n()) 

advice_avg <- study2_time2 %>%
  group_by(Condition, Advice) %>%
  summarize(n = dplyr::n()) %>%
  dplyr::mutate(conditionCount = ifelse(Condition == 'Likable',count[count$Condition == 'Likable',]$number,count[count$Condition == 'Accurate',]$number)) %>%
  dplyr::mutate(percent = n/conditionCount*100) %>% 
  dplyr::select(Condition,Advice,percent)
  

# Create a facet plot with condition as the facet variable and gender as the line color
ggplot(advice_avg, aes(x = Advice, y = percent, color = Condition)) +
  geom_line(linewidth = 0.75)+
  scale_color_manual(values = c("#FFA500", "#4169E1"),
                      labels = c("Accurate", "Likable")) +
  scale_y_continuous(name = "% of participants") +
  scale_x_continuous(name = "Advised Rank") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(1.2, "lines"),          
        legend.text = element_text(size = 12),
        plot.title = element_text(hjust = 0.5, size = 20))

```

Again, we found supportive evidence that males are more confident about
themselves, giving higher initial rank prediction (more attractive) than
female participants
(`r round(mean(study2_time1[study2_time1$gender == "Male",]$Guess),2)`
vs.
`r round(mean(study2_time1[study2_time1$gender == "Female",]$Guess),2)`,
`r papaja::apa_print(t.test(Guess ~ gender, data = study2_time1, var.equal = T))$statistic`).

Examining advice from Stage 2 participants, we found that participants
advised a lower rank (that is, more attractive) in the likeability
treatment than the accuracy treatment
(`r round(mean(study2_time2[study2_time2$Condition == "Likable",]$Advice),2)`
vs
`r round(mean(study2_time2[study2_time2$Condition == "Accurate",]$Advice),2)`,
`r papaja::apa_print(t.test(Advice ~ Condition, data = study2_time2, var.equal = T))$statistic`).
As seen in Figure \@ref(fig:study2advice), notably, however, even in the
accuracy treatment, participants advised a lower rank than they
themselves had provided
(`r papaja::apa_print(t.test(study2_time2[study2_time2$Condition == "Accurate",]$Advice, mu = 7, var.equal = T))$statistic`).
This suggests that advisers inferred that advising someone that they
were more attractive would make the adviser appear more likeable and
therefore provided advice that communicated a more favorable impression
of the participants' attractiveness.

```{r study2regs}

modelsummary::modelsummary(list(`(1)` = lm(likable ~ Advice, data = study2_time3),
                                `(2)` = lm(warmAvg ~ Advice, data = study2_time3),
                                `(3)` = lm(trustAvg ~ Advice, data = study2_time3)),
                           vcov = c("classical", "classical", "classical"),
                           coef_map = c("Advice" = "Advised Rank",
                                        "(Intercept)" = "Constant"),
                           stars = T,
                           gof_map = list(list('raw' = 'nobs', 'clean' = 'N', 'fmt' = 0)),
                           caption = "When individuals receive advice that implies a high level of attractiveness (lower rank), they tend to perceive the advice giver as more likable (Column 1), warm (Column 2). Column 3 shows that advisors are rated as more trustworthy when they advice lower ranks, but this relationship is only directional.") %>%
  kableExtra::kable_styling(latex_option = "scale_down")

```

```{r study2rating, fig.cap="The evaluation of advice givers"}

rating <- study2_time3 %>%
  group_by(Advice) %>%
  summarise(Likable = mean(likable), Warm = mean(warmAvg), Trust = mean(trustAvg))

ggplot(rating, aes(x = Advice)) +
  geom_line(aes(y = Likable, color = "Likable"), linewidth = 0.75) +
  geom_line(aes(y = Warm, color = "Warm"), linewidth = 0.75) +
  geom_line(aes(y = Trust, color = "Trust"), linewidth = 0.75) +
  scale_color_manual(values = c("Likable" = "#FFA500", "Warm" = "#4169E1", "Trust" = "grey"),
                      labels = c("Likable", "Warm", "Trust")) +
  scale_y_continuous(name = "Rating") +
  scale_x_continuous(name = "Advised Rank") +
  theme_bw() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "grey"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(1.2, "lines"),          
        legend.text = element_text(size = 12),
        plot.title = element_text(hjust = 0.5, size = 20))


```

To test whether flattering advice brings more positive evaluation of
advisers, we create a single index for warmth and trustworthiness by
taking average of ratings on warmth, friendliness, good-naturedness,
trustworthiness for the former and ratings on trustworthiness, and
sincerity for the later. As seen in Figure \@ref(fig:study2rating), we
found that advisees who suggested that the advisee was more attractive
(lower rank) were indeed rated as more likeable and warm (b = -0.109, p
\< .05; b = -0.133, p \< .001; See Columns 1 and 2 of Table 2).
Interestingly, these benefits are not at the cost of being hypocritical;
advisers are also seen as more trustworthy, but the result is only
directional (Column 3 of Table 2). Notably, however, we were not powered
to do a comparison across the two experimental groups.

# Discussion

dvice substantially shapes people's career and personal outcomes.
Therefore, understanding advisers' motivations has far-reaching
implications. We propose that advisers' focus on advisees' expectations
and the desire to avoid disappointing them can induce gender differences
in the advice given and ultimately in the outcomes achieved. Gender
differences in career outcomes contribute significantly to the gender
wage gap, with women often holding less senior positions and
experiencing worse career outcomes, even with equal performance.
Therefore, identifying a factor such as expectations that contributes to
this gap has important implications for organizational practice.

In this paper, we examine a novel and unexplored factor of
advice-giving: interpersonal considerations. Advisers who take into
account the advisee's belief utility inflate their advice to convey a
more favorable impression, leading to worse decisions and gender
differences. Men have more inflated expectations than women, and taking
expectations into account can lead to gender differences, even without
discrimination. We ruled out discrimination as a separate channel by
withholding information about the gender of the advisee. Finally, we
expect that advisers are aware of the tradeoff between giving accurate
advice and avoiding disappointment and, thus, are less likely to inflate
their advice when advisees can compensate them after observing the
outcome of their choice.

The findings of this study suggest that organizations can promote gender
equity by calibrating employees' expectations or shifting advisers'
attention away from them. While these measures may not entirely solve
the gender gap problem, they can make a meaningful difference when
unconscious bias interventions have failed to show any effect
[@Chang2019; @Paluck2009]. We believe our findings may be particularly
relevant to societies that value signaling in interpersonal
relationships, including Hong Kong and many Asian countries. Future work
could also examine whether people in different cultures place more
weight on expectations.

# References

```{=tex}
\begingroup
\singlespacing
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\noindent
```
::: {#refs}
:::

```{=tex}
\endgroup
```
